Simulation of Raman sideband cooling with large Lamb-Dicke factor and high initial temperature

### 2014-05-14 report/paper.tex
[report] results

\subsection{Calculations of the coupling coefficients and Evolution of the density matrix}
\begin{figure}
  \begin{center}
    \includegraphics[width=7cm]{../raman_0.8_1.png}
  \end{center}
  \caption{Coupling strength to lower vibrational state for different initial states.}
  \label{fig-raman-curve}
\end{figure}
\begin{figure}
  \begin{center}
    \includegraphics[width=7cm]{../pump_0.8_0_curve.png}
  \end{center}
  \caption{Possibilities of going in to other vibrational states (i.e. heating) during optical pumping for different initial level $n_0$. The pumping beam is sent in from the orthogonal direction.}
  \label{fig-pump-curve}
\end{figure}
In order to simulate the evolution of the system, it is necessary to calculate the strength of the sideband, i.e. the coupling between different vibrational levels. For the Raman transition, this can be directly calculated from equation \ref{lamb-dicke} and the result is shown in figure \ref{fig-raman-curve}. As we can see, due to the large Lamb-Dicke factor, there is significant coupling to lower vibrational state even for $\delta n$ as large as $8$.\\

@@

\section{Results of the simulation}
\begin{figure}
  \begin{center}
    \includegraphics[width=7cm]{../cool_process.png}
  \end{center}
  \caption{Distribution in vibrational levels (including both hyperfine ground states) at different time. States with $n>30$ are not shown in the figure.}
  \label{fig-cool-process}
\end{figure}
\begin{figure}
  \begin{center}
    \includegraphics[width=7cm]{../n_decrease.png}
  \end{center}
  \caption{Decrease of $n$ during the cooling process.}
  \label{fig-n-decrease}
\end{figure}

\section{Conclusion.}

>>>>

\begin{figure}
  \begin{center}
    \includegraphics[width=7cm]{../raman_0.8_1.png}
  \end{center}
  \caption{Coupling strength to lower vibrational state for different initial states.}
  \label{fig-raman-curve}
\end{figure}
\begin{figure}
  \begin{center}
    \includegraphics[width=7cm]{../pump_0.8_0_curve.png}
  \end{center}
  \caption{Possibilities of going in to other vibrational states (i.e. heating) during optical pumping for different initial level $n_0$. The pumping beam is sent in from the orthogonal direction.}
  \label{fig-pump-curve}
\end{figure}
\subsection{Calculations of the coupling coefficients}
In order to simulate the evolution of the system, it is necessary to calculate the strength of the sideband, i.e. the coupling between different vibrational levels. For the Raman transition, this can be directly calculated from equation \ref{lamb-dicke} and the result is shown in figure \ref{fig-raman-curve}. As we can see, due to the large Lamb-Dicke factor, there is significant coupling to lower vibrational state even for $\delta n$ as large as $8$.\\

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=7cm]{../n_decrease.png}
  \end{center}
  \caption{Decrease of $n$ during the cooling process.}
  \label{fig-n-decrease}
\end{figure}
\begin{figure}
  \begin{center}
    \includegraphics[width=7cm]{../cool_process.png}
  \end{center}
  \caption{Distribution in vibrational levels (including both hyperfine ground states) at different time. States with $n>30$ are not shown in the figure.}
  \label{fig-cool-process}
\end{figure}
\subsection{Evolution of the density matrix}
The evolution of the density matrix, including both the coherent Raman transition and the ``non-coherent'' optical pumping, is governed by the optical Bloch equation.
\eqar{
  \diff{\rho_{ij}}{t}=&-\frac12\sum_k\paren{\Gamma_{ki}+\Gamma_{kj}}\rho_{ij}+\ui\sum_k\paren{\rho_{ik}\Omega_{kj}-\rho_{kj}\Omega_{ik}}\\
  \diff{\rho_{ii}}{t}=&\sum_k\rho_{kk}\Gamma_{ik}-\sum_k\Gamma_{ki}\rho_{ii}+\ui\sum_k\paren{\rho_{ik}\Omega_{ki}-\rho_{ki}\Omega_{ik}}
}
where $\Gamma_{ij}$ and $\Omega_{ij}$ are the pumping rate and the Rabi frequency of the Raman transition between states $i$ and $j$ respectively. Since in the weak coupling limit, i.e. when the off resonance coupling is negligible, the only two time scales are the $\Gamma$'s and $\Omega$'s, we normalize the maximum Rabi frequency to $1$ in the simulation.

\section{Results of the simulation}
The simulation of the Raman sideband cooling is done by minimizing the average vibrational levels after a certain time. The free parameters in the simulation are the $\delta n$ for the Raman transition and the relative rate of the optical pumping $\Gamma/\Omega$. In order to get a stable and reliable simulation result, the change of these parameters are limited to smooth functions, which are further limited to exponential decays in this simulation for simplicity.\\

The result of this simple optimization is a sequence with $\delta n$ decreases from $12$ to $2$ and with the optical pumping rate decreases from $2$ to $1$. Figure \ref{fig-n-decrease} shows the change in the average vibrational levels during the cooling process. We can clearly see that even if the Lamb-Dicke factor is large and we started with a high tempeature, the Raman sideband cooling with $\delta n\gg1$ can still efficiently cool the atom in the trap to $12$ times colder then the initial temperature. Figure \ref{fig-cool-process} is the energy level distribution at different time which shows that more than $75\%$ of the atoms are in the two lowest vibrational levels at the end of the cooling.

\section{Conclusion}
In this paper, I have shown with numerical simulation of the optical Bloch equation that Raman sideband cooling with $\delta n\gg1$ can be used to effeciently cool neutral atoms in a optical dipole trap with $\eta\approx1$ and high initial temperature. Future improvement of this might include automatic optimization of the sequence and also the simulation of cooling in all three dimensions.
